### 2013-10-07 paper/old_stuff/rejected_text.tex
Intro, research questions, and outline

The complexity of a file or a system is not the same as the complexity of the
changes made to the system. For example, a change to the error message in an
otherwise complex file, represents a simple change in a complex file.

The goal of this work, is to develop measures of the complexity of a change.

A change to the eA change that modifies the output message shown to users

Traditional complexity measures often require the whole file or system and were
not designed to measure the complexity of code fragments \cite{Hindle2008ICPC}.
Further, multiple studies have found that these measures correlate very
strongly with the lines of code in a file
\cite{Graves2000IEEE,Leszak2002JSS,Herraiz2007MSR} and may not provide much
additional information about the complexity of the system.

System 
Simple changes can be made to complex files and complex changes can be made to simple files. For example, a one line ch
One
point to note is that